The Negative Effect of Traffic Noise on House Prices: A Landscape Hedonic Analysis

### 2014-07-07 reports/NoiseHedonicRaR.tex
added GCVsFromMCSim figure, added captions to remaining figures in paper

\begin{table}
\caption{Matrix of Pearson Correlation Coefficients for Quantitative Variables}\label{tab:cor}
%\includegraphics[width = \textwidth]{../graphs/CorrelationMatrix}
\begin{center}
\small
\tabcolsep=0.05cm
\makebox[\linewidth][c]{
\begin{tabular}{l|rrrrrrrrrrrrr}
 & \multicolumn{1}{x{.35in}}{Price} & \multicolumn{1}{x{.38in}}{House} & \multicolumn{1}{x{.35in}}{Lot} & 
 \multicolumn{1}{x{.35in}}{Built} & \multicolumn{1}{x{.35in}}{Noise} & \multicolumn{1}{x{.35in}}{CBD} & \multicolumn{1}{x{.35in}}{Park} & \multicolumn{1}{x{.35in}}{Lake} & \multicolumn{1}{x{.35in}}{Shop} & \multicolumn{1}{x{.35in}}{White} & 
 \multicolumn{1}{x{.35in}}{U18} & \multicolumn{1}{x{.35in}}{Inc.} & 
 \multicolumn{1}{x{.35in}}{Test} \tabularnewline 
  \hline
Sale Price & 1 &  &  &  &  &  &  &  &  &  &  &  &  \tabularnewline
House Size & 0.74 & 1 &  &  &  &  &  &  &  &  &  &  &  \tabularnewline
Lot Size  & 0.38 & 0.45 & 1 &  &  &  &  &  &  &  &  &  &  \tabularnewline
Year Built  & 0.45 & 0.52 & 0.47 & 1 &  &  &  &  &  &  &  &  &  \tabularnewline
Noise & -0.24 & -0.12 & -0.04 & -0.25 & 1 &  &  &  &  &  &  &  &  \tabularnewline
Distance to CBD  & 0.31 & 0.45 & 0.43 & 0.58 & -0.09 & 1 &  &  &  &  &  &  &  \tabularnewline
Distance to Park  & 0.25 & 0.28 & 0.22 & 0.42 & -0.17 & 0.57 & 1 &  &  &  &  &  &  \tabularnewline
Distance to Lake & -0.04 & -0.10 & -0.25 & -0.17 & -0.01 & -0.09 & 0.20 & 1 &  &  &  &  &  \tabularnewline
Distance to Shop  & 0.27 & 0.27 & 0.15 & 0.37 & -0.24 & 0.46 & 0.38 & 0.15 & 1 &  &  &  &  \tabularnewline
Percent White & 0.24 & 0.15 & 0.25 & 0.15 & -0.09 & 0.34 & 0.16 & -0.14 & 0.14 & 1 &  &  &  \tabularnewline
Percent Under 18 & 0.26 & 0.30 & 0.11 & 0.36 & -0.19 & 0.29 & 0.32 & 0.09 & 0.32 & -0.22 & 1 &  &  \\ 
Income & 0.52 & 0.50 & 0.41 & 0.57 & -0.24 & 0.53 & 0.38 & -0.12 & 0.36 & 0.35 & 0.32 & 1 &  \tabularnewline
Elem.\ Test Scores & 0.29 & 0.28 & 0.29 & 0.34 & -0.09 & 0.35 & 0.22 & -0.10 & 0.18 & 0.31 & 0.10 & 0.44 & 1 \tabularnewline 
\end{tabular} }
\end{center} 
\end{table}

@@

We first conducted a Monte Carlo simulation recommended in \citet{Fotheringham2002} to see if the improved model fit with our LWR model was due to random chance. The intuition of the test is as follows - if the improved model fit isn't significant, then we should be able to see similar improvements in model fit by using the LWR model with the same houses but in a different spatial arrangement. Failing to recreate a spatial arrangement that yields a similar model fit as to that obtained with our actual data is then evidence that our spatial arrangement and the improved model fit obtained from a spatially heterogeneous model is significant.

We conducted 100 different experiments in which we randomly swapped the locations of our house transactions. In each case we then estimated the LWR model using 10 different bandwidths ranging from the nearest 100 to 4,000 sales just as before and calculated the GCV score. Every single time the smallest Generalized Cross-Validation score was obtained at the largest bandwidth. That is, in 100 consecutive opportunities to fit our data at a local level of a couple hundred nearest sales up to a couple thousand nearby sales, the GCV score was always smallest for the largest possible bandwidth. In other words, after running a thousand different local regression models on the shuffled data, we never obtained model performance anywhere near the LWR model with our true locational data. This seems to be strong evidence that the increased ability to predict house prices with a local model is not due to chance and that there is spatial non-stationarity in the hedonic function.

@@

A few patterns are clear in Figures~\ref{fig:AcreBetas} -- \ref{fig:NoiseBetas}. For instance, across all four bandwidths we see that the hedonic\footnote{Remember that the dependent variable is the log of sales price.} coefficient on lot size (measured in acres) is largest near the central business district of St.\ Paul. Thus, consistent with standard urban economic theory, to the extent that the hedonic coefficient is correlated with the price of land, our results suggest the price of land is highest near the central business district and declines with distance from the CBD. This tendency is consistent across the four different bandwidths presented (other specific results available upon request). The LWR coefficients on house size (measured in thousands of square feet of finished space) also seem to exhibit spatial variation across the different bandwidths. In the west-central portion of our study area (a collection of notably desirable neighborhoods) we see a cluster of large LWR coefficients, while those areas in the southernmost portion of our region (suburban areas) have the smallest coefficients.

@@

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.1\textwidth]{../graphs/AcreBeta_allBandwidth}}
 \caption{Simula.}\label{fig:AcreBetas}
\end{figure}

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.1\textwidth]{../graphs/FinSqFtBeta_allBandwidth}}
 \caption{Simula.}\label{fig:HouseBetas}
\end{figure}

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.1\textwidth]{../graphs/NoiseBeta_allBandwidth}}
 \caption{Simula.}\label{fig:NoiseBetas}
\end{figure}

We now conduct a second series of simulation experiments aimed at testing whether the spatial variation exhibited in particular coefficients is significant or due to random chance. The mechanics of the second simulation are very similar to the those described in the previous section. We randomly change the spatial arrangment of our house sales and estimate the LWR model. However, in this simulation we restrict the model to small bandwidths and compare the spread of the regression coefficient estimates under the random arrangment of the data to the spread found in our original LWR models with the actual spatial arrangement. If an independent variable in our hedonic regression function exhibits a stationary relationship with sales prices, then we should see similar levels of dispersion in the randomly arranged data as in our true data. However, if the dispersion we see in our random data is \emph{smaller} than the spread we observed in our actual data, that is evidence consistent with the variable having a non-stationary relationship. That is, the variation in regression coefficients obtained in our data is higher than we would observe from random chance.

We present the results of this simulation for the same three different bandwidths in Table~\ref{tab:LWR}- 500, 650, and 1,000 nearest observations. Figures~\ref{fig:MCsds500} and \ref{fig:MCsds500} visualize the results of the second simulation for different bandwidths. Each sub-figure displays the distribution of LWR coefficient standard deviations obtained from the 100 iterations of the Monte Carlo simulation and the standard deviation obtained with the actual data (in red). Note that it is the case that the standard deviations obtained with the actual data are typically substantially larger than those obtained in the simulation for most of the variables. However, the variation obtained in the noise coefficients simulations is not dissimilar to our actual LWR results, suggesting that the spatial non-stationarity exhibited in this coefficient is perhaps due to chance and not true spatial non-stationarity in the noise coefficient.

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{../graphs/MCsimResultsSDs500}}
 \caption{Simulation Results for nearest 500 observations. Actual LWR Coefficient Standard Deviations \textcolor{red}{(red line)} vs.\ Distribution of Monte Carlo Simulation Standard Deviations by Variable. Figures in which the red line is to the right of the black distribution suggest spatial non-stationarity in effect of the variable on house prices in our study region.}\label{fig:MCsds500}
\end{figure}

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{../graphs/MCsimResultsSDs650}}
 \caption{Simulation Results for nearest 650 observations. Actual LWR Coefficient Standard Deviations \textcolor{red}{(red line)} vs.\ Distribution of Monte Carlo Simulation Standard Deviations by Variable. Figures in which the red line is to the right of the black distribution suggest spatial non-stationarity in effect of the variable on house prices in our study region.}\label{fig:MCsds650}
\end{figure}

[Insert figure with nearest 1,000 observations here]

@@

Previous research has suggested that the impact of noise can change with time and economic conditions. For instance, \citet{Wilhelmsson2000} found that the traffic noise penalty to be stronger in the 1990s near Stockholm, Sweden compared to the 1980s. \citet{Cohen2009} also found the (airport) noise penalty to be larger in the early 2000s compared to the late 1990s in Atlanta, Georgia. We know that house prices in the United States fell dramatically during and after the Great Recession of 2008-09. We know less about the patterns of change in the hedonic price functions during and after the fall in house prices. \citet{Cho2011b} found evidence from hedonic analysis in Nashville, Tennessee, that consumers' willingness to pay for environmental amenities (such as proximity to open space and water views) decreased during the recession as compared to the previous economic boom.

Figures~\ref{fig:AcreTime}--\ref{fig:NoiseTime} show the spatial distribution of some our LWR hedonic coefficients for bandwidths of 500 and 2,000 nearest sales within the last year for the sale years 2006--2010. Regardless of bandwidth, the figures show that the estimated LWR coefficients are different over time. In particular, the lot size and house size coefficients tend to decrease over time. What is surprising, however, is that the negative noise coefficient estimates tend to be more negative in later years of our study. Figure~\ref{fig:NoiseTime} shows that the spatial patterns are similar across years (the most negative coefficients tend to be in the same locations across years), but the average noise coefficient tends to be more negative for the later years.

\begin{sidewaysfigure}
 \includegraphics[trim = 0cm 2cm 0cm 0cm, clip = true, width = \textwidth]{../graphs/Acre_50_200_ByYear}
 \caption{Simula.}
 \label{fig:AcreTime}
\end{sidewaysfigure}

\begin{sidewaysfigure}
 \includegraphics[trim = 0cm 2cm 0cm 0cm, clip = true, width = \textwidth]{../graphs/FinFt_50_200_ByYear}
 \caption{Simula.}
 \label{fig:HouseTime}
\end{sidewaysfigure}

\begin{sidewaysfigure}
 \includegraphics[trim = 0cm 2cm 0cm 0cm, clip = true, width = \textwidth]{../graphs/Noise_50_200_ByYear}
 \caption{Simula.}
 \label{fig:NoiseTime}
\end{sidewaysfigure}

We also find statistical evidence consistent with the estimated noise coefficients being more negative in the later time periods. This finding is robust to multiple different formulations and specifications. Tables~\ref{tab:noisebetatime500} and \ref{tab:noisebetatime2000} present the results of regressions using our estimated LWR noise coefficient as the dependent variable and time and spatial characteristics as explanatory variables for LWR bandwidths of 500 and 2,000 nearest sales.

Both tables are structured similarly. Column (1) finds evidence of a negative linear trend in the LWR noise coefficients over time - predicting a LWR noise coefficient of approximately twice the magnitude for the end of our study period as for the beginning. Column (2) includes a quadratic term and again predicts noise coefficients much more negative towards the end of our study period compared to the beginning. We repeat the regressions from (1) and (2) in (3) and (4) but this time also include city fixed effects in order to control for any potential locational differences of house transactions over time. Again, we see very similar results - the noise coefficient is more negative in later time periods. We also limit our analysis to just the city of St.\ Paul and repeat the analysis from (1) and (2) in columns (5) and (6). Again, we estimate the noise coefficient to be more negative in later time periods.

% Requires LaTeX packages: dcolumn 
\begin{table}[!htbp] \centering 
  \label{tab:noisebetatime500} 
\scriptsize 
\begin{tabular}{@{\extracolsep{-1pt}}lD{.}{.}{6} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} } 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable: LWR Noise Coefficients (bandwidth = 500 nearest sales)}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)} & \multicolumn{1}{c}{(5)} & \multicolumn{1}{c}{(6)}\\ 
\hline \\[-1.8ex] 
Months since Jan 2005 & -0.00005^{***} & 0.00003^{***} & -0.00005^{***} & 0.00003^{***} & -0.0001^{***} & 0.00004^{***} \\ 
  & (0.000001) & (0.000004) & (0.000001) & (0.000004) & (0.000002) & (0.00001) \\ 
  & & & & & & \\ 
 (Months since Jan 2005)$^2$ &  & -0.000001^{***} &  & -0.000001^{***} &  & -0.000001^{***} \\ 
  &  & (0.00000005) &  & (0.00000004) &  & (0.0000001) \\ 
  & & & & & & \\ 
 Constant & -0.0016^{***} & -0.0028^{***} & -0.0021^{***} & -0.0033^{***} & -0.0015^{***} & -0.0032^{***} \\ 
  & (0.00003) & (0.0001) & (0.00003) & (0.0001) & (0.0001) & (0.0002) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
City Fixed Effects & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} \\ 
\hline \\[-1.8ex] 
Observations & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{8,003} & \multicolumn{1}{c}{8,003} \\ 
R$^{2}$ & \multicolumn{1}{c}{0.1273} & \multicolumn{1}{c}{0.1377} & \multicolumn{1}{c}{0.2931} & \multicolumn{1}{c}{0.3028} & \multicolumn{1}{c}{0.1566} & \multicolumn{1}{c}{0.1704} \\ 
\hline 
\hline \\[-1.8ex]  
\end{tabular} 
  \caption{LWR Noise Coefficient Over Time} 
\end{table}

>>>>

\begin{table}
\caption{Matrix of Pearson Correlation Coefficients for Quantitative Variables}\label{tab:cor}
\begin{center}
\small
\tabcolsep=0.05cm
\makebox[\linewidth][c]{
\begin{tabular}{l|rrrrrrrrrrrrr}
 & \multicolumn{1}{x{.35in}}{Price} & \multicolumn{1}{x{.38in}}{House} & \multicolumn{1}{x{.35in}}{Lot} & 
 \multicolumn{1}{x{.35in}}{Built} & \multicolumn{1}{x{.35in}}{Noise} & \multicolumn{1}{x{.35in}}{CBD} & \multicolumn{1}{x{.35in}}{Park} & \multicolumn{1}{x{.35in}}{Lake} & \multicolumn{1}{x{.35in}}{Shop} & \multicolumn{1}{x{.35in}}{White} & 
 \multicolumn{1}{x{.35in}}{U18} & \multicolumn{1}{x{.35in}}{Inc.} & 
 \multicolumn{1}{x{.35in}}{Test} \tabularnewline 
  \hline
Sale Price & 1 &  &  &  &  &  &  &  &  &  &  &  &  \tabularnewline
House Size & 0.74 & 1 &  &  &  &  &  &  &  &  &  &  &  \tabularnewline
Lot Size  & 0.38 & 0.45 & 1 &  &  &  &  &  &  &  &  &  &  \tabularnewline
Year Built  & 0.45 & 0.52 & 0.47 & 1 &  &  &  &  &  &  &  &  &  \tabularnewline
Noise & -0.24 & -0.12 & -0.04 & -0.25 & 1 &  &  &  &  &  &  &  &  \tabularnewline
Distance to CBD  & 0.31 & 0.45 & 0.43 & 0.58 & -0.09 & 1 &  &  &  &  &  &  &  \tabularnewline
Distance to Park  & 0.25 & 0.28 & 0.22 & 0.42 & -0.17 & 0.57 & 1 &  &  &  &  &  &  \tabularnewline
Distance to Lake & -0.04 & -0.10 & -0.25 & -0.17 & -0.01 & -0.09 & 0.20 & 1 &  &  &  &  &  \tabularnewline
Distance to Shop  & 0.27 & 0.27 & 0.15 & 0.37 & -0.24 & 0.46 & 0.38 & 0.15 & 1 &  &  &  &  \tabularnewline
Percent White & 0.24 & 0.15 & 0.25 & 0.15 & -0.09 & 0.34 & 0.16 & -0.14 & 0.14 & 1 &  &  &  \tabularnewline
Percent Under 18 & 0.26 & 0.30 & 0.11 & 0.36 & -0.19 & 0.29 & 0.32 & 0.09 & 0.32 & -0.22 & 1 &  &  \\ 
Income & 0.52 & 0.50 & 0.41 & 0.57 & -0.24 & 0.53 & 0.38 & -0.12 & 0.36 & 0.35 & 0.32 & 1 &  \tabularnewline
Elem.\ Test Scores & 0.29 & 0.28 & 0.29 & 0.34 & -0.09 & 0.35 & 0.22 & -0.10 & 0.18 & 0.31 & 0.10 & 0.44 & 1 \tabularnewline 
\end{tabular} }
\end{center} 
\end{table}

@@

We first conducted a Monte Carlo simulation recommended in \citet{Fotheringham2002} to see if the improved model fit with our LWR models was due to random chance. The intuition of the test is as follows - if the improved model fit isn't significant, then we should be able to see similar improvements in model fit by using the LWR model with the same houses but in a different spatial arrangement. Failing to recreate a spatial arrangement that yields a similar model fit as to that obtained with our actual data is then evidence that our spatial arrangement and the improved model fit obtained from a spatially heterogeneous model is significant.

We conducted 100 different experiments in which we randomly resampled (without replacement) the locations of our house transactions. In each case we then estimated the LWR model using 10 different bandwidths ranging from the nearest 100 to 4,000 sales just as before and calculated the GCV score. Figure~\ref{fig:GCVSIM} shows the distribution of the GCV scores we obtained for the simulated data for four selected bandwidths relative to the score we obtained with our actual data. Note that the GCV distributions for the simulated data a strongly related their bandwidth - smaller bandwidths have larger GCV scores. That is, with our randomized location data, the GCV score is inversely proportional to the bandwidth. Additionally, note that no GCV scores from the simulated data are near our GCV score with the actual data. In every single instance the smallest Generalized Cross-Validation score was obtained at the largest bandwidth. That is, in 100 consecutive opportunities to fit our data at a local level of a couple hundred nearest sales up to a couple thousand nearby sales, the GCV score was always smallest for the largest possible bandwidth. In other words, after running a thousand different local regression models on the shuffled data, we never obtained model performance anywhere near the LWR model with our true locational data. This seems to be strong evidence that the increased ability to predict house prices with a local model is not due to chance and that there is spatial non-stationarity in the hedonic function.

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.1\textwidth]{../graphs/GCVsFromMCsim}}
 \caption{This figure shows the distribution of GCV scores obtained from simulations with spatially randomized data and the GCV score from our data. Each solid curve represents the smoothed density plot of 100 different instances of the simulation. The dashed line suggests that our actual minimum GCV score was not obtained by random chance as no other GCV scores come close.}
 \label{fig:GCVSIM}
\end{figure}

@@

A few patterns are clear in Figures~\ref{fig:AcreBetas} -- \ref{fig:NoiseBetas}. For instance, across all four bandwidths we see that the hedonic coefficient\footnote{Remember that the dependent variable is the log of sales price.} on lot size (measured in acres) is largest near the central business district of St.\ Paul. Thus, consistent with standard urban economic theory, to the extent that the hedonic coefficient is correlated with the price of land, our results suggest the price of land is highest near the central business district and declines with distance from the CBD. This tendency is consistent across the four different bandwidths presented (other specific results available upon request). The LWR coefficients on house size (measured in thousands of square feet of finished space) also seem to exhibit spatial variation across the different bandwidths. In the west-central portion of our study area (a collection of notably desirable neighborhoods) we see a cluster of large LWR coefficients, while those areas in the southernmost portion of our region (suburban areas) have the smallest coefficients.

@@

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.1\textwidth]{../graphs/AcreBeta_allBandwidth}}
 \caption{This figure shows an interpolated surface (using Inverse Distance Weighting) for the estimated LWR coefficients (dependent variable = ln(sales price)) for the lot size variable (in acres) obtained across four different bandwidths: 500, 650, 1,000, and 2,000 nearest observations. In all cases there is a clear inverse relationship between the coefficient magnitudes and distance to the central business district.}
 \label{fig:AcreBetas}
\end{figure}

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.1\textwidth]{../graphs/FinSqFtBeta_allBandwidth}}
 \caption{This figure shows an interpolated surface (using Inverse Distance Weighting) for the estimated LWR coefficients (dependent variable = ln(sales price)) for the finished house size variable (in 1,000s of square feet) obtained across four different bandwidths: 500, 650, 1,000, and 2,000 nearest observations. In all cases some of the highest coefficient values are in the west-central portion of our study area and the lowest coefficients tend to be in the far south.}
 \label{fig:HouseBetas}
\end{figure}

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.1\textwidth]{../graphs/NoiseBeta_allBandwidth}}
 \caption{This figure shows an interpolated surface (using Inverse Distance Weighting) for the estimated LWR coefficients (dependent variable = ln(sales price)) for the noise variable (in dB) obtained across four different bandwidths: 500, 650, 1,000, and 2,000 nearest observations. In all cases some of the most negative coefficient values are in the west-central portion of our study area with the smallest in magnitude coefficients tending to be near the central business district.}
 \label{fig:NoiseBetas}
\end{figure}

We now conduct a second series of simulation experiments aimed at testing whether the spatial variation exhibited in particular coefficients is significant or due to random chance. The mechanics of the second simulation are very similar to the those described in the previous section. We randomly change the spatial arrangment of our house sales and estimate the LWR model. However, in this simulation we compare the \emph{spread} (rather than the degree of model fit) of the regression coefficient estimates under the random arrangment of the data to the spread found in our original LWR models with the actual spatial arrangement, conditional on bandwidth. If an explanatory variable in our hedonic regression function exhibits a stationary relationship with sales prices, then we should see similar levels of dispersion in the randomly arranged data as in our true data. However, if the dispersion we see in our randomized data is \emph{smaller} than the spread we observed in our actual data, that is evidence consistent with the variable having a non-stationary relationship. That is, the variation in regression coefficients obtained in our data is higher than we observe from random chance.

We conducted this simulation 100 times for the same four different bandwidths in Table~\ref{tab:LWR}- 500, 650, 1,000 and 2,000 nearest observations. Across all bandwidths we obtain similar results, so for brevity's sake we present the results for the 500 and 2,000 bandwidth sizes\footnote{Additional results available on request.} in Figure~\ref{fig:MCsim2}. The solid red (blue) lines show the distribution of the coefficient standard deviations we obtained from the 100 simulations for the 2,000 (500) nearest sale bandwidths. Across all variables it is the case the the distribution of standard deviations is significantly larger for the smaller bandwidth. That is, when we use smaller bandwidths we tend to see greater dispersion in regression coefficients. We also plot the standard deviation of our coefficients from our actual data as vertical dashed lines (again, color denotes the bandwidth size (red=2,000 and blue=500). For most variables, the standard deviation of the LWR coefficients obtained with our actual data is significantly greater than any standard deviation obtained with the randomized simulated data. That is, across our simulated data we never saw anywhere near the level of dispersion in regression coefficients for house size, lot size, etc. that we observe with our actual data. This seems to be strong evidence that many of our regression coefficients in our hedonic function exhibit spatial non-stationarity. However, for some variables (note the distributions for noise, owner occupancy and percent under age 18) for which our observed spread in the LWR coefficients is similar when we used the simulated data. This simulation does not present evidence that these variables exhibit a non-stationary relationship in our hedonic function.

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[height=.9\textheight]{../graphs/MCsimResultsSDsk500and2000final}}
 \caption{This figure shows the actual standard deviation in LWR coefficients as compared to those obtained from 100 different instances of spatial randomization. Results are presented for two different bandwidths (nearest 500 and 2,000) for our regression variables. In all cases the distribution of standard deviations for the smaller bandwidth is larger. In most cases the standard deviation of LWR coefficients for the actual data (the dashed line) is substantially higher than any standard deviation obtained with simulated data.}
 \label{fig:MCsim2}
\end{figure}

@@

We know that house prices in the United States fell dramatically during and after the Great Recession of 2008-09. We know less about the patterns of change in the hedonic price functions during and after the fall in house prices. \citet{Cho2011b} found evidence from hedonic analysis in Nashville, Tennessee, that consumers' willingness to pay for environmental amenities (such as proximity to open space and water views) decreased during the recession as compared to the previous economic boom. Previous research has suggested that the impact of noise can change with time and economic conditions. For instance, \citet{Wilhelmsson2000} found that the traffic noise penalty to be stronger in the 1990s near Stockholm, Sweden compared to the 1980s. \citet{Cohen2009} also found the (airport) noise penalty to be larger in the early 2000s compared to the late 1990s in Atlanta, Georgia.

Figures~\ref{fig:AcreTime}--\ref{fig:NoiseTime} show the spatial distribution of some our LWR hedonic coefficients for bandwidths of 500 and 2,000 nearest sales within the last year for the sale years 2006--2010. For both bandwidths presented\footnote{Additional results available upon request.}, the figures show that the estimated LWR coefficients are different over time. In particular, the lot size and house size coefficients tend to decrease over time. These results are consistent with standard economic theory, as wealth and income decrease, consumer willingness to pay for normal goods decreases. What is surprising, however, is that the negative noise coefficient estimates also tend to be more negative in later years of our study. Figure~\ref{fig:NoiseTime} shows that the spatial patterns are similar across years (the most negative coefficients tend to be in the same locations across years), but the average noise coefficient tends to be more negative for the later years.

\begin{sidewaysfigure}
 \includegraphics[trim = 0cm 2cm 0cm 0cm, clip = true, width = \textwidth]{../graphs/Acre_50_200_ByYear}
 \caption{This figure shows the interpolated LWR lot size coefficients across space and time for two bandwidths (nearest 500 and 2,000). Note that the spatial pattern obtained is similar across time periods, but the average value tends to decline over time.}
 \label{fig:AcreTime}
\end{sidewaysfigure}

\begin{sidewaysfigure}
 \includegraphics[trim = 0cm 2cm 0cm 0cm, clip = true, width = \textwidth]{../graphs/FinFt_50_200_ByYear}
 \caption{This figure shows the interpolated LWR finished house size coefficients across space and time for two bandwidths (nearest 500 and 2,000). Note that the spatial pattern obtained is similar across time periods, but the average value tends to decline over time.}
 \label{fig:HouseTime}
\end{sidewaysfigure}

\begin{sidewaysfigure}
 \includegraphics[trim = 0cm 2cm 0cm 0cm, clip = true, width = \textwidth]{../graphs/Noise_50_200_ByYear}
 \caption{This figure shows the interpolated LWR noise coefficients across space and time for two bandwidths (nearest 500 and 2,000). Note that the spatial pattern obtained is similar across time periods, but the average value tends to decline over time.}
 \label{fig:NoiseTime}
\end{sidewaysfigure}

We also find statistical evidence consistent with the estimated noise coefficients being more negative in the later time periods. This finding is robust to multiple different formulations and specifications. Tables~\ref{tab:noisebetatime500} and \ref{tab:noisebetatime2000} present the results of regressions using our estimated LWR noise coefficient as the dependent variable and time and spatial characteristics as explanatory variables for LWR bandwidths of 500 and 2,000 nearest sales. Both tables are structured similarly. Column (1) finds evidence of a negative linear trend in the LWR noise coefficients over time - predicting a LWR noise coefficient of approximately twice the magnitude for the end of our study period as for the beginning. Column (2) includes a quadratic term and again predicts noise coefficients much more negative towards the end of our study period compared to the beginning. We repeat the regressions from (1) and (2) in (3) and (4) but this time also include city fixed effects in order to control for any potential locational differences of house transactions over time. Again, we see very similar results - the noise coefficient is more negative in later time periods. We also limit our analysis to just the city of St.\ Paul and repeat the analysis from (1) and (2) in columns (5) and (6). Again, we estimate the noise coefficient to be substantially more negative in later time periods.

% Requires LaTeX packages: dcolumn 
\begin{table}[!htbp] \centering 
  \label{tab:noisebetatime500} 
  \caption{LWR Noise Coefficient Over Time (bandwidth = 500 nearest sales)} 
\scriptsize 
\begin{tabular}{@{\extracolsep{-1pt}}lD{.}{.}{6} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} } 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable: LWR Noise Coefficient (bandwidth = 500 nearest sales)}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)} & \multicolumn{1}{c}{(5)} & \multicolumn{1}{c}{(6)}\\ 
\hline \\[-1.8ex] 
Months since Jan 2005 & -0.00005^{***} & 0.00003^{***} & -0.00005^{***} & 0.00003^{***} & -0.0001^{***} & 0.00004^{***} \\ 
  & (0.000001) & (0.000004) & (0.000001) & (0.000004) & (0.000002) & (0.00001) \\ 
  & & & & & & \\ 
 (Months since Jan 2005)$^2$ &  & -0.000001^{***} &  & -0.000001^{***} &  & -0.000001^{***} \\ 
  &  & (0.00000005) &  & (0.00000004) &  & (0.0000001) \\ 
  & & & & & & \\ 
 Constant & -0.0016^{***} & -0.0028^{***} & -0.0021^{***} & -0.0033^{***} & -0.0015^{***} & -0.0032^{***} \\ 
  & (0.00003) & (0.0001) & (0.00003) & (0.0001) & (0.0001) & (0.0002) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
City Fixed Effects & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} \\ 
\hline \\[-1.8ex] 
Observations & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{8,003} & \multicolumn{1}{c}{8,003} \\ 
R$^{2}$ & \multicolumn{1}{c}{0.1273} & \multicolumn{1}{c}{0.1377} & \multicolumn{1}{c}{0.2931} & \multicolumn{1}{c}{0.3028} & \multicolumn{1}{c}{0.1566} & \multicolumn{1}{c}{0.1704} \\ 
\hline 
\hline \\[-1.8ex]  
\end{tabular} 
\end{table}

% Requires LaTeX packages: dcolumn 
\begin{table}[!htbp] \centering 
  \caption{LWR Noise Coefficient Over Time (bandwidth = 2,000 nearest sales)} 
  \label{tab:noisebetatime2000} 
\scriptsize 
\begin{tabular}{@{\extracolsep{-1pt}}lD{.}{.}{6} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} } 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable: LWR Noise Coefficient (bandwidth = 2,000 nearest sales)}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)} & \multicolumn{1}{c}{(5)} & \multicolumn{1}{c}{(6)}\\ 
\hline \\[-1.8ex] 
Months since Jan 2005 & -0.00005^{***} & 0.00003^{***} & -0.00005^{***} & 0.00003^{***} & -0.00005^{***} & 0.0001^{***} \\ 
  & (0.0000003) & (0.000002) & (0.0000003) & (0.000002) & (0.000001) & (0.000004) \\ 
  & & & & & & \\ 
 (Months since Jan 2005)$^2$ &  & -0.000001^{***} &  & -0.000001^{***} &  & -0.000001^{***} \\ 
  &  & (0.00000002) &  & (0.00000002) &  & (0.0000001) \\ 
  & & & & & & \\ 
 Constant & -0.0014^{***} & -0.0026^{***} & -0.0016^{***} & -0.0028^{***} & -0.0016^{***} & -0.0033^{***} \\ 
  & (0.00001) & (0.00003) & (0.00002) & (0.00003) & (0.00003) & (0.0001) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
City Fixed Effects & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} \\ 
\hline \\[-1.8ex] 
Observations & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{8,003} & \multicolumn{1}{c}{8,003} \\ 
R$^{2}$ & \multicolumn{1}{c}{0.3828} & \multicolumn{1}{c}{0.4142} & \multicolumn{1}{c}{0.5128} & \multicolumn{1}{c}{0.5435} & \multicolumn{1}{c}{0.3454} & \multicolumn{1}{c}{0.3988} \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}
